---
title: 'XDMOD '
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2023-09-25"
---

# Prerequisits

This example use python API for XDMoD analytic framework.
So you already need to have installed python version. 
In R you should install `reticulate` library and bound the python with XDMoD framework.

## Quick Conda Install

Here is a quick tip on installation (tested in WSL ubuntu)

```bash
conda install requests
conda create -n xdmod-notebooks -y python=3.11 r=4.3
conda activate xdmod-notebooks
# base-notebook
conda install -y 'jupyterlab' 'notebook' 'jupyterhub' 'nbclassic'
# scipy-notebook
conda install -y  'altair' 'beautifulsoup4' 'bokeh' 'bottleneck' 'cloudpickle' \
    'conda-forge::blas=*=openblas' \
    'cython' 'dask' 'dill' 'h5py' 'ipympl' 'ipywidgets' 'jupyterlab-git' \
    'matplotlib-base' 'numba' 'numexpr' 'openpyxl' 'pandas' 'patsy' 'protobuf' \
    'pytables' 'scikit-image' 'scikit-learn' 'scipy' 'seaborn' 'sqlalchemy' \
    'statsmodels' 'sympy' 'widgetsnbextension' 'xlrd'
# r-notebook
conda install -y 'r-base' 'r-caret' 'r-crayon' 'r-devtools' 'r-e1071' \
    'r-forecast' 'r-hexbin' 'r-htmltools' 'r-htmlwidgets' 'r-irkernel' \
    'r-nycflights13' 'r-randomforest' 'r-rcurl' 'r-rmarkdown' 'r-rodbc' \
    'r-rsqlite' 'r-shiny' 'r-tidymodels' 'r-tidyverse' 'unixodbc'

# Other
conda install -y 'pymysql' 'requests' \
    'r-plotly' 'r-repr' 'r-irdisplay' 'r-pbdzmq' 'r-reticulate' 'r-cowplot' \
    'r-rjson' 'r-dotenv'
    
# Install xdmod-data
pip install --upgrade 'xdmod-data>=1.0.0,<2.0.0' python-dotenv tabulate

# Install rstudio server
wget -q "https://download2.rstudio.org/server/jammy/amd64/rstudio-server-2023.12.1-402-amd64.deb"
dpkg -i rstudio-server-*-amd64.deb
rm rstudio-server-*-amd64.deb

# specify which version of r to use
echo "rsession-which-r=$(which R)" | sudo tee -a /etc/rstudio/rserver.conf
```

Launch rserver at <http://localhost:8787>


# Library Load

```{r setup, include=T}
# load libraries
knitr::opts_chunk$set(echo = TRUE)

library(dotenv)
library(reticulate)
library(tidyverse)
library(plotly)
library(knitr)
library(rmarkdown)

# replace xdmod-notebooks with conda enviroment for python to use
use_condaenv("xdmod-notebooks")
# on windows ~ points to Documents, adjust accordently
load_dot_env(path.expand("~/xdmod-data.env"))
```


# Use XDMoD Analytical Framework as Python API

## Using python chunks

First, initialize the XDMoD Data Warehouse.
Run the code below to prepare for getting data from the XDMoD data warehouse at the given URL.

```{python}
# This is a python chunk
# Initialize the XDMoD Data Warehouse
from xdmod_data.warehouse import DataWarehouse
dw = DataWarehouse('https://xdmod.access-ci.org')
```

Next, get the data.
Run the code below to use the `get_data()` method to request data from XDMoD and load them into a DataFrame. This example gets the number of active users of ACCESS-allocated resources over a 4-month period. Each of the parameters of the method will be explained later in this notebook. Use `with` to create a runtime context; this is also explained later in this notebook.

```{python}
# This is a python chunk
# Get data
with dw:
    data = dw.get_data(
        duration=('2023-01-01', '2023-04-30'),
        realm='Jobs',
        metric='Number of Users: Active',
    )
```

Slight modification to converted DataFrame:
`Tidyverse` approach do not use row-names and convert text to date.

```{r}
# Use data in R
df <- py$data %>% 
    rownames_to_column(var="Time") %>% # Move row names to column
    tibble() %>% # use newer data.frame
    mutate(Time=ymd(Time)) # convert character to date

df %>% paged_table()
```

Plot the data

```{r}
# plot
ggplotly(
    ggplot(df, aes(x=Time,y=`Number of Users: Active`)) +
        geom_line()
)
```

## Calling from R using reticulate

Now we can repeat same things using `reticulate` generated R interface,
the steps are same: init warehouse, get-data, convert data and use it.

```{r}
# Initialize the XDMoD Data Warehouse
warehouse <- import("xdmod_data.warehouse")
dw <- warehouse$DataWarehouse('https://xdmod.access-ci.org')
```

```{r}
# Query data
with(dw,{
    data2 <- dw$get_data(
        duration=c('2023-01-01', '2023-04-30'),
        realm='Jobs',
        metric='Number of Users: Active',
)})
```

```{r rows.print=20}
# Use data in R
df2 <-data2 %>% 
    rownames_to_column(var="Time") %>% # Move row names to column
    tibble() %>% # use newer data.frame
    mutate(Time=ymd(Time)) # convert character to date

df2
```

```{r}
# plot
ggplotly(
    ggplot(df2, aes(x=Time,y=`Number of Users: Active`)) +
        geom_line()
)
```

## Do further data processing
You can do further processing on the DataFrame to produce analysis and plots beyond those that are available in the XDMoD portal.

Run the code below to add a column for the day of the week:

```{r}
df2$weekday <- factor(
    weekdays(df2$Time), 
    levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
             "Saturday", "Sunday"))
df2
```

Run the code below to show a box plot of the data grouped by day of the week:

```{r}
ggplotly(
    ggplot(df2, aes(x=weekday,y=`Number of Users: Active`)) +
        geom_boxplot()
)
```

# Details of the `get_data()` method
Now that you have seen a basic example of using the `get_data()` method, read below for more details on how it works.

## Wrap data warehouse calls in a runtime context
XDMoD data is accessed over a network connection, which involves establishing connections and creating temporary resources. To ensure these connections and resources are cleaned up properly in spite of any runtime errors, you should call data warehouse methods within a **runtime context** by using Python's `with` statement to wrap the execution of XDMoD queries, store the result, and execute any long running calculations outside of the runtime context, as in the template below. `reticulate` package of R reflect this layout.

```r
with(dw, {
    # XDMoD queries would go here
    pass
    })
# Data processing would go here
pass
```

## Default parameters

The `get_data()` method has a number of parameters; their default values are shown below, and the parameters are explained in more detail further below.

```{r}
with(dw, {
    data <- dw$get_data(
        duration='Previous month',
        realm='Jobs',
        metric='CPU Hours: Total',
        dimension='None',
        filters=list(),
        dataset_type='timeseries',
        aggregation_unit='Auto'
    )})
data
```

## Duration

The **duration** provides the time constraints of the data to be fetched from the XDMoD data warehouse.

As already seen, you can specify the duration as start and end times:

```{r}
with(dw, {
    data <- dw$get_data(duration=c('2023-01-01', '2023-04-30'))})
data
```

You can instead specify the duration using a special string value; a list of the valid values can be obtained by calling the `get_durations()` method.

```{r}
with(dw, {
    durations = dw$get_durations()
    })
durations
```

## Realm

A **realm** is a category of data in the XDMoD data warehouse.
You can use the `describe_realms()` method to get a DataFrame containing the list of available realms.


```{r}
with(dw, {
    realms <- dw$describe_realms()
    })
realms
```

## Metric

A **metric** is a statistic for which data exists in a given realm. You can use the `describe_metrics(realm)` method to get a DataFrame containing the list of valid metrics in the given realm. The realm must be passed in as a string.

```{r}
with(dw,{
    metrics <- dw$describe_metrics('Jobs')})
metrics %>% kable(format="html")
```
## Dimension
A **dimension** is a grouping of data. You can use the `describe_dimensions(realm)` method to get a DataFrame containing the list of valid dimensions in the given realm. The realm must be passed in as a string.

```{r}
with(dw,{
    dimensions <- dw$describe_dimensions('Jobs') })
dimensions %>% kable(format="html")
```

## Pass in realms, metrics, and dimensions using labels or IDs

For methods in the API that take realms, metrics, and/or dimensions as arguments, you can pass them in as their labels or their IDs.

```{r}
with(dw, {
    data <- dw$get_data(
        duration='10 year',
        realm='Allocations',
        metric='NUs: Allocated', # 'allocated_nu' also works
        dimension='Resource Type',  # 'resource_type' also works
    )})
data
```

## Filters

**Filters** allow you to include only data that have certain values for given dimensions. You can use the `get_filter_values(realm, dimension)` method to get a DataFrame containing the list of valid filter values for the given dimension in the given realm. The realm and dimension must be passed in as strings.

```{r}
with(dw, {
    filter_values <- dw$get_filter_values('Jobs', 'Resource') }) # 'resource' also works
filter_values
```

For methods in the API that take filters as arguments, you must specify the filters as a dictionary in which the keys are dimensions (labels or IDs) and the values are string filter values (labels or IDs) or sequences of string filter values. For example, to return only data for which the field of science is biophysics and the resource is either NCSA Delta GPU or TACC Stampede2:

```{r}
with(dw,{
    data <- dw$get_data(
        filters=list(
            'Field of Science'='Biophysics', # 'fieldofscience': '246' also works
            'Resource'= c( # 'resource' also works
                'NCSA DELTA GPU', # '3032' also works
                'STAMPEDE2 TACC' # '2825' also works
            )
        ),
    )})

data
```

## Dataset Type
The **dataset type** can either be 'timeseries' (the default), in which data is grouped by a time [aggregation unit](#Aggregation-unit), or 'aggregate', in which the data is aggregated across the entire [duration](#Duration). For 'aggregate', the results are returned as a Pandas Series rather than a DataFrame.

```{r}
with(dw, {
    data <- dw$get_data(dataset_type='timeseries')})

data
```

### Aggregation unit

The **aggregation unit** specifies how data is aggregated by time. You can get a list of valid aggregation units by calling the `get_aggregation_units()` method.

```{r}
with(dw,{
    aggregation_units <- dw$get_aggregation_units()})

aggregation_units
```


## Additional Examples
For additional examples, please see the [xdmod-notebooks repository](https://github.com/ubccr/xdmod-notebooks).



